---
title: "Math 536 Final Exam: Appendix R Code"
author: "Shamir Naran"
date: "Due May 15, 2020"
output: pdf_document  
---


## Problem 1


```{r, message = FALSE}
library(ggplot2)
library(esquisse)
```


```{r}
# Read in the data

baseball <- read.csv("baseball.csv", header = TRUE)
```


```{r}
# Data Cleaning

baseball <- na.omit(baseball) # remove all rows containing an NA
baseball <- baseball[-c(1,4)] # we don't need "X" or the "pitcher" column
```


```{r}
# change called ball to 0 and called strike to 1

baseball$description <- ifelse(baseball$description == "ball", 0, 1)
```


```{r}
# Define Strike Zone (from the internet)

TopStrikeZone <- 3.5
BotStrikeZone <- 1.5
LeftStrikeZone <- -0.85
RightStrikeZone <- 0.85
StrikeZone <- data.frame(
  x=c(LeftStrikeZone, LeftStrikeZone, RightStrikeZone, RightStrikeZone, LeftStrikeZone),
  y=c(BotStrikeZone, TopStrikeZone, TopStrikeZone, BotStrikeZone, BotStrikeZone))
```


```{r message=FALSE, warning=FALSE}
ggplot() +
  geom_point(data = data.frame(baseball$plate_x[baseball$description == 0],baseball$plate_z[baseball$description == 0]),
             mapping = aes(baseball$plate_x[baseball$description == 0],baseball$plate_z[baseball$description == 0]),
             col = "gray50", shape = 1) +
  geom_point(data = data.frame(baseball$plate_x[baseball$description == 1],baseball$plate_z[baseball$description == 1]),
             mapping = aes(baseball$plate_x[baseball$description == 1],baseball$plate_z[baseball$description == 1]),
             col = "blue2", shape = 3) +
  geom_path(aes(x,y), data = StrikeZone, lwd = 1,col = "black") +
  xlim(-5, 5) + ylim(-0.5, 5) +
  labs(title = "Scatterplot of Balls and Strikes", x = "Horizontal Location (ft.)", y = "Vertical Location (ft.)")
  
```


```{r}
# Identify "Fringe Pitches" (add and subtract 0.05 ft. from the strike zone)

baseball = baseball[!(baseball$plate_x > -0.75 & baseball$plate_x < 0.75 & baseball$plate_z > 1.7 & baseball$plate_z < 3.2),]
baseball = baseball[(baseball$plate_x > -1 & baseball$plate_x < 1 & baseball$plate_z > 1.4 & baseball$plate_z < 3.5),]

```


```{r message=FALSE, warning=FALSE}
ggplot() +
  geom_point(data = data.frame(baseball$plate_x[baseball$description == 0],baseball$plate_z[baseball$description == 0]),
             mapping = aes(baseball$plate_x[baseball$description == 0],baseball$plate_z[baseball$description == 0]),
             col = "gray50", shape = 1) +
  geom_point(data = data.frame(baseball$plate_x[baseball$description == 1],baseball$plate_z[baseball$description == 1]),
             mapping = aes(baseball$plate_x[baseball$description == 1],baseball$plate_z[baseball$description == 1]),
             col = "blue2", shape = 3) +
  xlim(-1, 1) + ylim(1, 3.5) +
  labs(title = "Scatterplot of Fringe Pitches", x = "Horizontal Location (ft.)", y = "Vertical Location (ft.)")
```


```{r}
# Good mix of balls and strikes

length(which(baseball$description == 0)) # number of balls in data
length(which(baseball$description == 1)) # number of strikes in data

length(which(baseball$description == 0)) / length(baseball$description) # % of balls
length(which(baseball$description == 1)) / length(baseball$description) # % of strikes

```


```{r}
# Create new Variables (count, run differential, distance)

baseball[ , "count"] = ""
baseball[ , "run_diff"] = ""
baseball[ , "distance"] = ""

baseball[baseball$ball == 3 & baseball$strikes < 2,]$count = "3less2"
baseball[baseball$ball == 3 & baseball$strikes == 2,]$count = "full"
baseball[baseball$ball != 3,]$count = "other"

baseball$run_diff <- abs(baseball$home_score - baseball$away_score)

baseball$distance <- sqrt((baseball$plate_x)^2+(baseball$plate_z-2.4)^2)
```


```{r}
ggplot(baseball) +
  aes(x = plate_x, y = plate_z, colour = count) +
  geom_point(size = 1L) +
  scale_color_hue() +
  theme_minimal() +
  facet_wrap(vars(description)) +
  labs(title = "Cases", x = "Horizontal Location (ft.)", y = "Vertical Location (ft.)")
```


```{r}
# Fit a Logistic Regression Model

model <- glm(description ~ pitch_type + release_speed + balls + strikes + count + run_diff + distance,
             data = baseball, family = "binomial")
sum.model <- summary(model)
sum.model
```


```{r}
# Our diagnostic plot looks good

pearson <- (baseball$description - model$fit) / (model$fit * (1-model$fit))
est.ln.odds <- log(model$fit/(1-model$fit))
loess1 = loess(pearson~est.ln.odds)
plot(est.ln.odds,pearson)
lines(est.ln.odds[order(est.ln.odds)],loess1$fit[order(est.ln.odds)])
```


```{r}
# 95% Confidence Interval for all predictors

lb <- exp(sum.model$coefficients[,1] - qnorm(0.975) * sum.model$coefficients[,2])
ub <- exp(sum.model$coefficients[,1] + qnorm(0.975) * sum.model$coefficients[,2])

cbind(lb,ub)
```


Significant predictors are release speed, balls, strikes, and distance.


```{r}

# Some calculations

log(.99708)

log(9.458974e-01)
log(9.933937e-01)

log(5.922246e-01)
log(7.416091e-01)

log(4.766717e-04)
log(1.494886e-03)
```


According to this model, we're 95% confident that on average, holding all else constant ... 

a one unit increase in release speed will decrease the odds of a called strike between [5.56% and 0.663%].

a one unit increase in strikes will decrease the odds of a called strike between [52.38% and 52.38%].

a one unit increase in balls will increase the odds of a called strike between [5.62% ,32.08%].

a one unit increase in distance will decrease the odds of a called strike between [764.87% and 650.57%].


```{r}
# Lets predict holding all else constant

full <- data.frame(pitch_type = "FF", release_speed = 94.3, balls = 3, strikes = 2,
                   distance = 1, count = "full", run_diff = 1)

three.less.two <- data.frame(pitch_type = "FF", release_speed = 94.3, balls = 3, strikes = 1,
                   distance = 1, count = "3less2", run_diff = 1)

other <- data.frame(pitch_type = "FF", release_speed = 94.3, balls = 0, strikes = 0,
                   distance = 1, count = "other", run_diff = 1)


predict.full <- predict.glm(model, newdata = full, type = "response")
predict.three.less.two <- predict.glm(model, newdata = three.less.two, type = "response")
predict.other <- predict.glm(model, newdata = other, type = "response")

list(predict.full = predict.full, predict.three.less.two = predict.three.less.two, predict.other = predict.other)

```
